The value of intermediation

### 2022-09-01 src/paper/paper.tex
Minor formatting changes

\title{The value of intermediation -- results so far}
\author{Martin Stancsics}

@@

Consider a cooperative game with two types of players: a major player $P$ and $n$ smaller players. I.e., $N = \{P, A_1, \dots, A_n\}$. Assume that (1) no coalition of players can achieve a positive value without the participation of $P$ and (2) players $A_i$ are identical. Let $n_T(S)$ denote the number of players of type $T \in \{P, A\}$ in coalition $S$. Then, the value that any coalition $S \subset N$ can achieve is the following:
\begin{align*}
    v(S) = \begin{cases}
        0                              & \text{if } P \notin S \\
        f\left(\frac{n_A(S)}{n}\right) & \text{otherwise}.
    \end{cases}
\end{align*}
One can think of this setting as a (one-sided) platform and a number of potential entrants, if the entrants can only reach their prospective consumers through the platform.\footnote{While these assumptions might seem somewhat restrictive, they are relatively simple to relax.} First, let us characterize the conditions under which this game is superadditive. These features provide some theoretical support for using Shapley-values as bargaining outcomes, as papers creating a link between noncooperative bargaining and cooperative values often rely on such assumptions.\footnote{For example, the results in \textcite{gul1989bargaining} rely on superadditivity, while monotonicity suffices for \textcite[]{hart1996bargaining}.}

@@

While the previous procedure might look somewhat peculiar at first glance, its elements would not be out of place in more traditional (noncooperative) models, either. Principally, the two main ideas it relies are the following. First, bargaining outcomes should not depend on which player actually obtains the money, all that matters is the total that a given set of players can obtain.\footnote{For example, imagine the following two situations with two players. In the first one, the first player obtains a playoff of two dollars if they both decide to cooperate and nothing otherwise. In the second scenario, both players receive one dollar in the case of cooperation. If they can transfer monay between each other, bargaining outcomes should be the same in both scenarios.} Second, players during the bargaining foresee the (unique) subgame perfect equilibrium in the following subgame, and negotiate based on them

@@

Next steps:
\begin{itemize}
    \item Show how the equilibrium number of entrants depends on $N_P$
    \item Show how consumer welfare (which is proportional to $\log(N_f + N_P)$) depends on $N_P$
    \item Try to characterize the profit-maximizing $N_P$ and how it depends on the parameters
\end{itemize}

Why is it interesting?
\begin{itemize}
    \item In \textcite{anderson2021hybrid}, in the absence of a revenue-based fee, consumer welfare does not depend on the number/measure of the platform's products. (The idea is that a higher $N_P$ is offset by a lower $N_f$ and vice versa.) Consumer welfare effects materialize through the platform setting a different sales fee for different values of $N_P$.
    \item In this model, however, $N_P$ also changes the bargaining power of the platform, which in turn changes the profits obtained by the fringe. Therefore, $N_P$ does have an effect on the equilibrium measure of entrants and consumer welfare, even without revenue-based fees.
    \item Therefore, a paltform selling its own products is not only worriesome because of the ususal factors (possibility of steering, perverse incentives for setting sales fees, etc.), but also because it increases its bargaining power against the possible entrants.
\end{itemize}

@@

\begin{proof}[Proof of Lemma \ref{lem:entry_distr}] \textcolor{red}{(Might have to fix sum indices)} \\
    The probability of $P$ having at most fraction $t$ of the other players before itself is
    \begin{align*}
        \Pr(X_n \leq nt) &= \Pr(X_n \leq nt ) \\
        &= \prod_{j = nt + 1}^n \frac{j}{j + \lambda} \\
        &= \exp \Bigg( \underbrace{\sum_{j = nt + 1}^n \log(j) - \log(j+\lambda)}_{\equiv S_n} \Bigg).
    \end{align*}
    Taking limits,
    \begin{align*}
        \lim_{n \to \infty} S_n &= \lim_{n \to \infty} \sum_{j = nt + 1}^n \log(j) - \log(j+\lambda) \\
        &= \lim_{n \to \infty} \sum_{i = 1}^{n - nt} \log(nt + i) - \log(nt + i + \lambda) \\
        &= \lim_{n \to \infty} \frac{1}{n - nt} \sum_{i = 1}^{n - nt} \frac{\log \left( t + \frac{i}{n - nt} \right) - \log \left( t + \frac{i}{n - nt} + \frac{\lambda}{n - nt} \right)}{1 / (n - nt)}
    \end{align*}
    Let
    \begin{align*}
        \Delta_n(s) = \frac{\log \left( s \right) - \log \left( s + \frac{\lambda}{n - nt} \right)}{1 / (n - nt)}
    \end{align*}
    By lemma \ref{lemma:log_convergence}, 
    \begin{align*}
        \Delta_n \xrightarrow[]{\mathrm{u}} \lambda \frac{\mathrm{d}}{\mathrm{d}s}log(s) = -\frac{\lambda}{s}
    \end{align*}
    on the compact interval $[t, 1]$ for any $t > 0$ ($\xrightarrow[]{\mathrm{u}}$ denotes uniform convergence).

@@

    For $t=0$, simply observe that \textcolor{red}{Have to elaborate on this}
    \begin{align*}
        \lim_{n \to \infty} \prod_{j = 1}^n \frac{j}{j + \lambda} = 0 = 0^\lambda.
    \end{align*}
\end{proof}

@@

\begin{proof}[Proof of Corollary \ref{cor:multiple_platforms_2}] \textcolor{red}{(Assuming $f$ is continuously differentiable -- will have to relax this)}
    $f'(t)$ is non-negative, so $[1 - (1-t)^m] f'(t)$ is increasing in $m$ $\forall t \in [0, 1]$. As a consequence, $\varphi_A^{n, \infty}$ is also increasing in $m$ if $f'(t)$ is positive on some interval or, $f(t)$ is not constant. Conversely, $\varphi_{P}^{\infty, m} = 1 - \varphi_{A}^{\infty, m}$ is decreasing in $m$.
\end{proof}

>>>>

\title{The value of intermediation}
\author{Martin Stancsics}

@@

Consider a cooperative game with two types of players: a major player $P$ and $n$ smaller players. I.e., $N = \{P, A_1, \dots, A_n\}$. Assume that (1) no coalition of players can achieve a positive value without the participation of $P$ and (2) players $A_i$ are identical. Let $n_T(S)$ denote the number of players of type $T \in \{P, A\}$ in coalition $S$. Then, the value that any coalition $S \subset N$ can achieve is the following:
\begin{align*}
    v(S) = \begin{cases}
        0                              & \text{if } P \notin S \\
        f\left(\frac{n_A(S)}{n}\right) & \text{otherwise}.
    \end{cases}
\end{align*}
One can think of this setting as a (one-sided) platform and a number of potential entrants, if the entrants can only reach their prospective consumers through the platform.\footnote{While these assumptions might seem somewhat restrictive, they are relatively simple to relax.} First, let us characterize the conditions under which this game is superadditive. These features provide some theoretical support for using Shapley-values as bargaining outcomes, as papers creating a link between non-cooperative bargaining and cooperative values often rely on such assumptions.\footnote{For example, the results in \textcite{gul1989bargaining} rely on superadditivity, while monotonicity suffices for \textcite[]{hart1996bargaining}.}

@@

While the previous procedure might look somewhat peculiar at first glance, its elements would not be out of place in more traditional (non-cooperative) models, either. Principally, the two main ideas it relies are the following. First, bargaining outcomes should not depend on which player actually obtains the money, all that matters is the total that a given set of players can obtain.\footnote{For example, imagine the following two situations with two players. In the first one, the first player obtains a playoff of two dollars if they both decide to cooperate and nothing otherwise. In the second scenario, both players receive one dollar in the case of cooperation. If they can transfer monay between each other, bargaining outcomes should be the same in both scenarios.} Second, players during the bargaining foresee the (unique) subgame perfect equilibrium in the following subgame, and negotiate based on them.

@@

This paper proposed a new framework to think about the profit allocation in situations concerning a few large and a continuum of smaller players. The results are especially relevant for upstream-downstream situations, platform settings, and two-sided markets. The outcomes of the Shapley-value-based gain sharing rule generally concur with intuitive expectations, and the resulting formulas are more tractable than most explicit non-cooperative approaches.

The illustrative hybrid platform model also highlights the strengths of this approach. Despite its simplicity, it highlights a novel channel through which platforms selling their own products might lead to welfare losses: hybrid platforms have more bargaining power against potential entrants than traditional ones. This mechanism should probably also be considered when competition authorities assess mergers between platforms and contents providers.

There is a variety of ways through which this paper could be expanded or built upon. On the theoretical side, some of the stricter assumptions could be relaxed, such as fringe agents achieving a zero payoff without the platform. Besides, the convergence result for the weighted values could be strengthened if one could manage to show that any sequence along ever more refined partitions of the atomic players converges to the same limit \parencite[à la][]{fogelman1980asymptotic}. The possibilities for practical applications are no less interesting, either. As a first step, the majorly simplified example application could be expanded upon, so that it is a better description of certain real-life paltform markets. Finally, the cooperative approach to profit distribution could become an alternative for the more usual, but in some sense more extreme assumptions about the bargaining process widely used in industrial organization models.

% Next steps:
% \begin{itemize}
%     \item Show how the equilibrium number of entrants depends on $N_P$
%     \item Show how consumer welfare (which is proportional to $\log(N_f + N_P)$) depends on $N_P$
%     \item Try to characterize the profit-maximizing $N_P$ and how it depends on the parameters
% \end{itemize}

@@

\begin{proof}[Proof of Lemma \ref{lem:entry_distr}] %\textcolor{red}{(Might have to fix sum indices)} \\
    The probability of $P$ having at most fraction $t$ of the other players before itself is
    \begin{align*}
        \Pr(X_n \leq nt) &= \Pr(X_n \leq nt ) \\
        &= \prod_{j = nt + 1}^n \frac{j}{j + \lambda} \\
        &= \exp \Bigg( \underbrace{\sum_{j = nt + 1}^n \log(j) - \log(j+\lambda)}_{\equiv S_n} \Bigg).
    \end{align*}
    Taking limits,
    \begin{align*}
        \lim_{n \to \infty} S_n &= \lim_{n \to \infty} \sum_{j = nt + 1}^n \log(j) - \log(j+\lambda) \\
        &= \lim_{n \to \infty} \sum_{i = 1}^{n - nt} \log(nt + i) - \log(nt + i + \lambda) \\
        &= \lim_{n \to \infty} \frac{1}{n - nt} \sum_{i = 1}^{n - nt} \frac{\log \left( t + \frac{i}{n - nt} \right) - \log \left( t + \frac{i}{n - nt} + \frac{\lambda}{n - nt} \right)}{1 / (n - nt)}
    \end{align*}
    Let
    \begin{align*}
        \Delta_n(s) = \frac{\log \left( s \right) - \log \left( s + \frac{\lambda}{n - nt} \right)}{1 / (n - nt)}
    \end{align*}
    By lemma \ref{lemma:log_convergence}, 
    \begin{align*}
        \Delta_n \xrightarrow[]{\mathrm{u}} \lambda \frac{\mathrm{d}}{\mathrm{d}s}log(s) = -\frac{\lambda}{s}
    \end{align*}
    on the compact interval $[t, 1]$ for any $t > 0$ ($\xrightarrow[]{\mathrm{u}}$ denotes uniform convergence).

@@

    For $t=0$, simply observe that %\textcolor{red}{Have to elaborate on this}
    \begin{align*}
        \lim_{n \to \infty} \prod_{j = 1}^n \frac{j}{j + \lambda} = 0 = 0^\lambda.
    \end{align*}
\end{proof}

@@

\begin{proof}[Proof of Corollary \ref{cor:multiple_platforms_2}] (Assuming $f$ is continuously differentiable) %\textcolor{red}{(Assuming $f$ is continuously differentiable -- will have to relax this)}
    $f'(t)$ is non-negative, so $[1 - (1-t)^m] f'(t)$ is increasing in $m$ $\forall t \in [0, 1]$. As a consequence, $\varphi_A^{n, \infty}$ is also increasing in $m$ if $f'(t)$ is positive on some interval or, $f(t)$ is not constant. Conversely, $\varphi_{P}^{\infty, m} = 1 - \varphi_{A}^{\infty, m}$ is decreasing in $m$.
\end{proof}
